---
title: "data.processing"
author: "CJ Robinson"
date: "5/20/2019"
output: html_document
---
##Set Up

Set working directory to the location of the master replication folder

```{r setup}
knitr::opts_knit$set(root.dir = "")
```

Load required packages

```{r packages}
library(plyr)
library(dplyr)
library(ggplot2)
library(stargazer)
```

Load data from CSV file attained from GSS Data Explorer (https://gssdataexplorer.norc.org/)

```{r load_data}
survey_master <- read.csv("./data/survey_master.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

##Data Cleaning

Rename variables for better use in cleaning and analysis

```{r rename}
gss <- survey_master %>%
  rename(work = "Labor.force.status",
         needs_access = "Need.but.unable.to.use.public.transport.",
         findjob = "How.likely.r.make.effort.for.new.job.next.year",
         mobility = "Could.r.find.equally.good.job",
         jobsecurity = "Is.r.likely.to.lose.job",
         citysize = "Expanded.norc..size.code",
         income = "Respondents.income",
         race = "Race.of.respondent",
         gender = "Respondents.sex",
         educ = "Highest.year.of.school.completed",
         age = "Age.of.respondent",
         children = "Number.of.children",
         has_car = "Do.not.have.a.car",
         gov = "Govt.or.private.employee",
         prestige = 'Rs.occupational.prestige.score..2010.')
```

Filter for those not in the labor force (Unemployed or some form of employment)

```{r labor_force}
gss <- gss %>%
  filter(work != "School", 
         work != "Other", 
         work != "Keeping house",
         work != "Retired")
```

Filter for the four largest city sizes

```{r citysize}
gss <- gss %>%
  filter(citysize == "City gt 250000" |
           citysize == "City,50-250000" |
           citysize == "Suburb, lrg city" |
           citysize == "Suburb, med city")
```

Filter incomes under $60,000

```{r income}
gss <- gss %>%
  filter(income == "Under $1 000" ||
           income == "$1 000 to 2 999" |
           income == "$3 000 to 3 999" |
           income == "$4 000 to 4 999" |
           income == "$5 000 to 5 999" |
           income == "$6 000 to 6 999" |
           income == "$7 000 to 7 999" |
           income == "$8 000 to 9 999" |
           income == "$10000 to 12499" |
           income == "$12500 to 14999" |
           income == "$15000 to 17499" |
           income == "$17500 to 19999" |
           income == "$20000 to 22499" |
           income == "$22500 to 24999" |
           income == "$25000 to 29999" |
           income == "$30000 to 34999" |
           income == "$35000 to 39999" |
           income == "$40000 to 49999" |
           income == "$50000 to 59999")
```


## Dataset 1 - Employment Status

Selects for variables necessary to Dataset 1

```{r select1}
gss1 <- gss %>%
  select(work,
         needs_access,
         has_car,
         citysize,
         income,
         race,
         gender,
         children,
         age, 
         educ)
```

Filters out blank responses

```{r blanks1}
gss1[gss1 == "Not applicable"] <- NA
gss1[gss1 == "No answer"] <- NA
gss1[gss1 == "Dont know"] <- NA
gss1[gss1 == "Don't know"] <- NA
gss1[gss1 == "Dk na"] <- NA
gss1 <- na.omit(gss1)
```


Utilize the next section as a sort of "codebook" for values in the first cleaned dataset. This also converts the dataset to all numeric values. I assigned 0's to "No" and 1's to "Yes" when appropriate, and otherwise the coding decisions are arbitrary. 

Note: has_car is coded backward so values go negative (No car is a 0) to positive (car is a 1)
because the question asks the respondents if they DO NOT have a car

```{r coding1}
gss1$work[gss1$work == "Unempl, laid off"] <- 0
gss1$work[gss1$work == "Working fulltime"] <- 1
gss1$work[gss1$work == "Working parttime"] <- 1
gss1$work[gss1$work == "Temp not working"] <- 1

gss1$needs_access[gss1$needs_access == "No"] <- 0
gss1$needs_access[gss1$needs_access == "Yes"] <- 1

gss1$citysize[gss1$citysize == "Suburb, med city"] <- 0
gss1$citysize[gss1$citysize == "Suburb, lrg city"] <- 1
gss1$citysize[gss1$citysize == "City,50-250000"] <- 2
gss1$citysize[gss1$citysize == "City gt 250000"] <- 3

gss1$income[gss1$income == "Under $1 000"] <- 5000
gss1$income[gss1$income == "$1 000 to 2 999"] <- 1500
gss1$income[gss1$income == "$3 000 to 3 999"] <- 3500
gss1$income[gss1$income == "$4 000 to 4 999"] <- 4500
gss1$income[gss1$income == "$5 000 to 5 999"] <- 5500
gss1$income[gss1$income == "$6 000 to 6 999"] <- 6500
gss1$income[gss1$income == "$7 000 to 7 999"] <- 7500
gss1$income[gss1$income == "$8 000 to 9 999"] <- 9000
gss1$income[gss1$income == "$10000 to 12499"] <- 11250
gss1$income[gss1$income == "$12500 to 14999"] <- 13750
gss1$income[gss1$income == "$15000 to 17499"] <- 16250
gss1$income[gss1$income == "$17500 to 19999"] <- 18750
gss1$income[gss1$income == "$20000 to 22499"] <- 21250
gss1$income[gss1$income == "$22500 to 24999"] <- 23750
gss1$income[gss1$income == "$25000 to 29999"] <- 27500
gss1$income[gss1$income == "$30000 to 34999"] <- 32500
gss1$income[gss1$income == "$35000 to 39999"] <- 37500
gss1$income[gss1$income == "$40000 to 49999"] <- 45000
gss1$income[gss1$income == "$50000 to 59999"] <- 55000

gss1$race[gss1$race == "White"] <- 0
gss1$race[gss1$race == "Black"] <- 1
gss1$race[gss1$race == "Other"] <- 1

gss1$gender[gss1$gender == "Male"] <- 0
gss1$gender[gss1$gender == "Female"] <- 1

gss1$has_car[gss1$has_car == "No"] <- 1
gss1$has_car[gss1$has_car == "Yes"] <- 0

gss1$children[gss1$children == "Eight or more"] <- 8

gss1$children <- as.numeric(gss1$children)
```

Export Dataset 1 to a CSV file

```{r export1}
write.csv(gss1, "./data/gss1.csv")
```

## Dataset 2 - Employment Mobility

Selects for variables necessary to Dataset 2

```{r select2}
gss2 <- gss %>%
  select(work,
         needs_access,
         has_car,
         mobility,
         citysize,
         income,
         race,
         gender,
         children,
         age,
         educ)
```

Filters out blank responses

```{r blanks2}
gss2[gss2 == "Not applicable"] <- NA
gss2[gss2 == "No answer"] <- NA
gss2[gss2 == "Dont know"] <- NA
gss2[gss2 == "Don't know"] <- NA
gss2[gss2 == "Dk na"] <- NA
gss2 <- na.omit(gss2)
```

The main difference between these two coding sets is in reference to the variable "work." Dataset 1 includes unemployed respondents while Dataset 2 does not. 

```{r coding2}
gss2[gss2 == "Unempl, laid off"] <- NA
gss2 <- na.omit(gss2)

gss2$work[gss2$work == "Working fulltime"] <- 2
gss2$work[gss2$work == "Working parttime"] <- 1
gss2$work[gss2$work == "Temp not working"] <- 0

gss2$needs_access[gss2$needs_access == "No"] <- 0
gss2$needs_access[gss2$needs_access == "Yes"] <- 1

gss2$mobility[gss2$mobility == "Not easy"] <- 0
gss2$mobility[gss2$mobility == "Somewhat easy"] <- 1
gss2$mobility[gss2$mobility == "Very easy"] <- 2

gss2$citysize[gss2$citysize == "Suburb, med city"] <- 0
gss2$citysize[gss2$citysize == "Suburb, lrg city"] <- 1
gss2$citysize[gss2$citysize == "City,50-250000"] <- 2
gss2$citysize[gss2$citysize == "City gt 250000"] <- 3

gss2$income[gss2$income == "Under $1 000"] <- 5000
gss2$income[gss2$income == "$1 000 to 2 999"] <- 1500
gss2$income[gss2$income == "$3 000 to 3 999"] <- 3500
gss2$income[gss2$income == "$4 000 to 4 999"] <- 4500
gss2$income[gss2$income == "$5 000 to 5 999"] <- 5500
gss2$income[gss2$income == "$6 000 to 6 999"] <- 6500
gss2$income[gss2$income == "$7 000 to 7 999"] <- 7500
gss2$income[gss2$income == "$8 000 to 9 999"] <- 9000
gss2$income[gss2$income == "$10000 to 12499"] <- 11250
gss2$income[gss2$income == "$12500 to 14999"] <- 13750
gss2$income[gss2$income == "$15000 to 17499"] <- 16250
gss2$income[gss2$income == "$17500 to 19999"] <- 18750
gss2$income[gss2$income == "$20000 to 22499"] <- 21250
gss2$income[gss2$income == "$22500 to 24999"] <- 23750
gss2$income[gss2$income == "$25000 to 29999"] <- 27500
gss2$income[gss2$income == "$30000 to 34999"] <- 32500
gss2$income[gss2$income == "$35000 to 39999"] <- 37500
gss2$income[gss2$income == "$40000 to 49999"] <- 45000
gss2$income[gss2$income == "$50000 to 59999"] <- 55000

gss2$race[gss2$race == "White"] <- 0
gss2$race[gss2$race == "Black"] <- 1
gss2$race[gss2$race == "Other"] <- 1

gss2$gender[gss2$gender == "Male"] <- 0
gss2$gender[gss2$gender == "Female"] <- 1

gss2$has_car[gss2$has_car == "No"] <- 1
gss2$has_car[gss2$has_car == "Yes"] <- 0

gss2$children[gss2$children == "Eight or more"] <- 8
```

Export Dataset 2

```{r export1}
write.csv(gss2, "./data/gss2.csv")
```






