---
title: "main.analysis"
author: "CJ Robinson"
date: "5/22/2019"
output: html_document
---
## Set Up

Set working directory to the location of the master replication folder

```{r setup}
knitr::opts_knit$set(root.dir = "")
```

Load required packages

```{r packages}
library(plyr)
library(dplyr)
library(ggplot2)
library(stargazer)
```

Load the two datasets

```{r load_data}
gss1 <- read.csv(file = "./data/gss1.csv")
gss2 <- read.csv(file = "./data/gss2.csv")
```

Factor several variables for easier analysis in regression models. Factoring variables is important in doing regression analyses when there are non-linear differences in the changes of variables. For example, you cannot define a constant difference between not working and working parttime and working parttime and working fulltime. In this case, we factor the variable and R will create dummy variables for each value while leaving one out. 

After I factored the variables, I releveled them to change while variable is omitted from analysis. For the most part, I did this to omit the response with the least sample size. 

I then changed the numeric variables back into strings for easier visualization in the regression tables. 

```{r factor1}
gss1 <- gss1 %>%
    mutate(citysize = factor(gss1$citysize))

gss1$citysize <- relevel(gss1$citysize, 1)

gss1$citysize <- revalue(gss1$citysize, c("3" = "mostbig", 
                                        "2" = "largermedium", 
                                        "1" = "smallermedium", 
                                        "0" = "mostsmall"))

```

```{r factor2}
gss2 <- gss2 %>%
  mutate(citysize = factor(gss2$citysize),
         work = factor(gss2$work))

gss2$citysize <- relevel(gss2$citysize, 1)

gss2$work <- relevel(gss2$work, 1)

gss2$citysize <- revalue(gss2$citysize, c("3" = "mostbig", 
                                        "2" = "largermedium", 
                                        "1" = "smallermedium", 
                                        "0" = "mostsmall"))

gss2$work <- revalue(gss2$work, c("0" = "notworking", 
                                "1" = "parttime", 
                                "2" = "fulltime"))

```


## Bivariate Models - Dataset 1

Dataset 1, Car Ownership, Bivariate Logit Regression with just work and car owernship

```{r bivar_model_car1}
bivar_model_car1 <- glm(work ~ has_car, gss1, family = "binomial")
summary(bivar_model_car1)
```

Saves Model
```{r save_bivar_car1}
save(bivar_model_car1, file = "./data/bivar_model_car1.RData")
```


Dataset 2, Public Transit, Bivariate Logit Regression with just work and public transit

```{r bivar_model_pt1}
bivar_model_pt1 <- glm(work ~ needs_access, gss1, family = "binomial")
summary(bivar_model_pt1)
```

Saves Model
```{r save_bivar_pt1}
save(bivar_model_pt1, file = "./data/bivar_model_pt1.RData")
```

## Bivariate Models - Dataset 2

Dataset 2, Car Ownership, Bivariate OLS Regression with just mobility and car ownership

```{r bivar_model_car2}
bivar_model_car2 <- lm(mobility ~ has_car, gss2)
summary(bivar_model_car2)
```

Saves Model
```{r save_bivar_car2}
save(bivar_model_car2, file = "./data/bivar_model_car2.RData")
```

Dataset 2, Public Transit, Bivariate OLS Regression with just mobility and public transit

```{r bivar_model_pt2}
bivar_model_pt2 <- lm(mobility ~ needs_access, gss2)
summary(bivar_model_pt2)
```

Saves Model

```{r save_bivar_pt2}
save(bivar_model_pt2, file = "./data/bivar_model_pt2.RData")
```

## Multivariate Models - Dataset 1

Dataset 1, Car Ownership, Multivariate Logit Regression with controls

```{r multivar_model_car1}
multivar_model_car1 <- glm(work ~ has_car + 
                             income + 
                             age + 
                             citysize + 
                             race + 
                             gender + 
                             educ +
                             children, gss1, family = "binomial")
summary(multivar_model_car1)
```

Saves Model

```{r save_multivar_car1}
save(multivar_model_car1, file = "./data/multivar_model_car1.RData")
```

Dataset 1, Public Transit, Multivariate Logit Regression with controls

```{r multivar_model_pt1}
multivar_model_pt1 <- glm(work ~ needs_access + 
                             income + 
                             age + 
                             citysize + 
                             race + 
                             gender + 
                             educ +
                             children, gss1, family = "binomial")
summary(multivar_model_pt1)
```

Saves Model

```{r save_multivar_pt1}
save(multivar_model_pt1, file = "./data/multivar_model_pt1.RData")
```

## Multivariate Models - Dataset 2

Dataset 2, Car Ownership, Multivariate OLS Regression with controls

```{r multivar_model_car2}
multivar_model_car2 <- lm(mobility ~ has_car + 
                            citysize + 
                            income + 
                            race +
                            gender + 
                            educ + 
                            age + 
                            children + 
                            work, gss2)
summary(multivar_model_car2)
```

Saves Model

```{r save_multivar_car2}
save(multivar_model_car2, file = "./data/multivar_model_car2.RData")
```

Dataset 2, Public Transit, Multivariate OLS Regression with Controls

```{r multivar_model_pt2}
multivar_model_pt2 <- lm(mobility ~ needs_access + 
                            citysize + 
                            income + 
                            race +
                            gender + 
                            educ + 
                            age + 
                            children + 
                            work, gss2)
summary(multivar_model_pt2)
```

Saves Model

```{r save_multivar_pt2}
save(multivar_model_pt2, file = "./data/multivar_model_pt2.RData")
```



