---
title: "tables"
author: "CJ Robinson"
date: "5/23/2019"
output: html_document
---
## Set Up

Set working directory to the location of the master replication folder

```{r setup}
knitr::opts_knit$set(root.dir = "")
```

Load required packages

```{r packages}
library(plyr)
library(dplyr)
library(ggplot2)
library(stargazer)
```

Load the two datasets and regression analysis

```{r load_data}
gss1 <- read.csv(file = "./data/gss1.csv")
gss2 <- read.csv(file = "./data/gss2.csv")
load("./data/bivar_model_pt1.RData")
load("./data/bivar_model_pt2.RData")
load("./data/bivar_model_car1.RData")
load("./data/bivar_model_car2.RData")
load("./data/multivar_model_pt1.RData")
load("./data/multivar_model_pt2.RData")
load("./data/multivar_model_car1.RData")
load("./data/multivar_model_car2.RData")
```

Creates table for summary statistics for control variables in Dataset 1

```{r summary_1}
summarystat1 <- gss1 %>% 
  select(-X) %>%
  select(-has_car) %>%
  select(-needs_access)

writeLines(capture.output(stargazer(summarystat1, title = "Summary Statistics for Controls in Dataset 1", type = "latex")), "./tables/summary_1.tex")
```

Creates table for summary statistics for control variables in Dataset 2

```{r summary_2}
summarystat2 <- gss2 %>% 
  select(-has_car) %>%
  select(-needs_access) %>%
  select(-mobility)
  
writeLines(capture.output(stargazer(summarystat2, title = "Summary Statistics for Controls in Dataset 2", type = "latex")), "./tables/summary_2.tex")
```

Creates correlation table for Datset 1

```{r correlation_1}
cor1 <- gss1 %>% 
  select(-X) %>%
  select(-has_car) %>%
  select(-needs_access)

gss_cor1 <- cor(cor1)

writeLines(capture.output(stargazer(gss_cor1, title = "Correlation Matrix for Controls in Dataset 1",  type = "latex")), "./tables/correlation_1.tex")
```

Creates correlation table for Datset 2

```{r correlation_2}
cor2 <- gss2 %>% 
  select(-X) %>%
  select(-mobility) %>%
  select(-has_car) %>%
  select(-needs_access)

gss_cor2 <- cor(cor2)

writeLines(capture.output(stargazer(gss_cor2, title = "Correlation Matrix for Controls in Dataset 2",  type = "latex")), "./tables/correlation_2.tex")
```

Further examines race relationship in Dataset 2

```{r race_2}
race_analysis <- lm(race ~ mobility, gss2)

summary(race_analysis)

writeLines(capture.output(stargazer(race_analysis,
          title = "Bivariate Model of Race in Dataset 2",
          dep.var.labels = "Employment Mobility of Respondent",
          covariate.labels = "Race",
          type = "latex", align = TRUE, no.space = TRUE)), "./tables/race_2.tex")
```

Creates Bivariate Table for Dataset 1

```{r bivar_table_1}
writeLines(capture.output(stargazer(bivar_model_pt1, bivar_model_car1,
          title = "Bivariate Models in Dataset 1",
          dep.var.labels = "Employment Status of Respondent",
          covariate.labels = c("Needs Access to Public Transit",
                               "Has a car"),
          type = "latex", align = TRUE, no.space = TRUE)), "./tables/bivar_table_1.tex")
```

Creates Multivariate Table for Dataset 1

```{r multivar_table_1}
writeLines(capture.output(stargazer(multivar_model_pt1, multivar_model_car1,
          title = "Model Comparisons of Logit Regression in Dataset 1",
          dep.var.labels = "Employment Status of Respondent",
          covariate.labels = c("Needs Access to Public Transit",
                               "Has a Car",
                               "Income",
                               "Age",
                               "3rd LargestCity Size",
                               "2nd Largest City Size",
                               "1st Largest City Size",
                               "Race",
                               "Gender",
                               "Education",
                               "Children"),
          type = "latex", align = TRUE, no.space = TRUE)), "./tables/multivar_table_1.tex")

```

Creates Bivariate Table for Dataset 2

```{r bivar_table_2}
writeLines(capture.output(stargazer(bivar_model_pt2, bivar_model_car2,
          title = "Bivariate Models in Dataset 2",
          dep.var.labels = "Employment Mobility of Respondent",
          covariate.labels = c("Needs Access to Public Transit",
                               "Has a car"),
          type = "latex", align = TRUE, no.space = TRUE)), "./tables/bivar_table_2.tex")
```

Creates Multivariate Table for Dataset 2

```{r multivar_table_2}
writeLines(capture.output(stargazer(multivar_model_pt2, multivar_model_car2,
          title = "Model Comparisons of Different Variable Formulations in Dataset 2",
          dep.var.labels = "Employment Mobility of Respondent",
          covariate.labels = c("Needs Access to Public Transit",
                               "Has a Car",
                               "3rd Largest City Size",
                               "2nd Largest City Size",
                               "1st Largest City Size",
                               "Income",
                               "Race",
                               "Gender",
                               "Education",
                               "Age",
                               "Children",
                               "Part-Time Workers",
                               "Full-Time Workers"),
          type = "latex", align = TRUE, no.space = TRUE)), "./tables/multivar_table_2.tex")
```

